---
title: |
  | \hspace{5cm} The Past Decade in Music
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, warning=FALSE, message=FALSE}
#plotting and wrangling
library(tidyverse)
library(highcharter)
library(factoextra)
library(patchwork)
library(sf)

#ml/modeling
library(caret)

#working with time variables
library(lubridate)

#setting the seed for reproducability
set.seed(543)
```


```{r Importing Data, warning=FALSE, message=FALSE}
#data is just imported here. Script for preprocessing and other wrangling and 
#imports from the Spotify API located in this folder under
#Spotify_get_script.R
billboard <- read_csv("data/spotify_full_100_10s.csv") %>% 
  rename(id = id...12)
```


```{r Viewing the data}
billboard %>% 
  distinct(id, .keep_all = TRUE) %>% 
  group_by(album.name) %>% 
  hchart(type = "scatter", hcaes(energy, valence, group = album.name)) %>% 
  hc_legend(verticalAlign = "left", layout = "horizontal", x = 30, y = 15, fontSize = "10px") %>% 
  hc_title(text = "Happiness vs. Energeticness of the Top 100 Albums", align = "center")
```

```{r Spatial Visualization, message=FALSE, warning=FALSE}
#Getting the country codes
countries <- read_csv("data/country_codes.csv")


#Converting to a shape Object
new_data <- billboard %>% 
  left_join(y = countries, by = "country") %>% 
  group_by(album.name, country) %>% 
  summarise_if(is.numeric, mean) %>% 
  ungroup() %>% 
  st_as_sf(coords = c("Latitude", "Longitude"))

#Plotting countries in the world
hcmap(map = "custom/world-robinson-lowres",
      data = new_data,
      name = "Happiness of Music (Valence)",
      value = "valence",
      borderWidth = 0,
      nullColor = "#d3d3d3") %>%
  hc_colorAxis(
    stops = color_stops(colors = viridisLite::magma(10, begin = 0.1)),
    type = "logarithmic"
    ) %>% 
  hc_title(text = "Happiness of the Billboard Top 100 Albums Available Around the World")
```

### What is the relationship between happiness and the other variables? 

From the Spotify Dataset, there are a few parameters that are computed as functions of other parameters found in the data. These parameters are energy, loudness, and acousticness. For example, *loudness* is used in the calculations of *energy*. For this reason, these terms will be interaction terms in the regression model. 

```{r GLM}
#selecting all of the numeric attributes for a linear model
model_ready <- billboard %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select_if(is.numeric) %>%
  select(-c("tag_freq"))

# Setting up cross validation
ctrl <- trainControl(method = "cv", number = 10)

# trying to model the relationship between happiness and the other variables
lm_spotify <- train(valence ~ danceability + energy * loudness * acousticness + speechiness + 
                     tempo + instrumentalness + liveness + key, data = model_ready,
                    method = "lm",
                    preProc = c("scale", "center"),
                    trControl = ctrl)
```

```{r}
summary(lm_spotify$finalModel)
```

Let's take a look at these plots.

```{r}
#getting the years of release for the plots
billboard_year <- billboard %>% 
  mutate(year = lubridate::year(release_date)) %>% 
  distinct(id, .keep_all = TRUE)

# plotting individual relationships as a scatter plot
hchart(billboard_year, "scatter", hcaes(x = danceability, y = valence, group = year)) %>% 
  hc_title(text = "Valence and Danceability")
hchart(billboard_year, "scatter", hcaes(x = energy, y = valence, group = year)) %>% 
  hc_title(text = "Valence and Energy")
hchart(billboard_year, "scatter", hcaes(x = speechiness, y = valence, group = year)) %>% 
  hc_title(text = "Valence and Energy")
```

Lets see how the model fits the data

```{r, fig.align="center"}
#predicting the values of valence with the models
predictions <- predict(lm_spotify, billboard)

ggplot(billboard, aes(x = valence, y = predictions)) +
  geom_point(color= "orange") +
  geom_smooth(method = "lm") +
  theme_minimal() +
  labs(x = "Valence", y = "Predictions", title = "Predictions vs. Observations of Happiness in Music")
```



